---
title: "assignment04"
output: html_document
author: "author: Jordan Grover, jg2058"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


```

## Instructions

Your analysis should have four data visualizations of distinct graph types, made with ggplot2. Across all
four graphs, use a total of:
• Six different aesthetics (i.e. set inside aes());
• Six different non-aesthetic options (i.e. set outside aes());
• Five different geoms;
• Two different scales (meaning change the default scale used for at least two aesthetics).
Further, each graph must include:
• Correct usage of all visual encodings;
• Appropriate data sourcing (hint, check out the labs() function);
• Proper labeling of all visual encodings;
• An appropriate title and subtitle;

## Data Source

Pandemic Oversight Coronavirus Relief Fund details
https://www.pandemicoversight.gov/data-interactive-tools/interactive-dashboards/data-exports


##Visualization 1

Description of Viz 1: The goal of this visualization is to first help us understand who is actually receiving awards. The award recipients aren't clearly specified, so we will have to give them an extra category for each of the typical federal grant recipients: counties, cities, states, and Native American tribes. With that understanding, we can layer on top, how many awards they receive, how much, and for what purposes. 

```{r}
library(tidyverse)
library(ggplot2)
library(urbnthemes)
data <- read.csv("CRF_details.csv")

#List of Prime Recipients
#unique(data$Prime.recipient)


data %>%
   mutate(Type = if_else(str_detect(Prime.recipient,"STATE|COMMONWEALTH|GOVERNOR"),"State", if_else(str_detect(Prime.recipient,"COUNTY"),"County",if_else(str_detect(Prime.recipient,"TRIBE|NATION|TRIBAL|NATIVE|INDIAN|COUNCIL|COMMUNITY|RANCHERIA"), "Tribe",if_else(str_detect(Prime.recipient,"CITY|MUNICIPALITY|VILLAGE|TOWN|PUBELO"), "City, Town, or Village","Other")))))%>%
  group_by(Type)%>%
  summarize(award_count = length(unique(Prime.recipient)))%>%
  ggplot()+
  geom_col(mapping = aes(x=Type, y = award_count, fill = Type))+
  scale_fill_brewer(palette="palette_urbn")+
    labs(y = "Number of Recipients")+
  labs(title = "The vast majority of award recipients are tribal nations",subtitle = "Number of recipients by recipient type",caption = "Source: Pandemicoversight.gov")+
  geom_text(mapping = aes(x = Type, y = award_count, label = award_count),position = position_nudge(y = +20))

```

## Visualization 2

Now knowing the breakdown of recipient types, how many grants are they receiving, and for how much? Each award is broken down into a sub recipient, an award, and then a spending category. I would like to just count how many unique awards there are per prime recipient, and what the sum of their awards are. To do that, I will make a new data set with Prime Recipient, Distinct Award, and Award Amount. There is no distinct ID for prime award, so I have made one by combining Prime Recipient with Award Amount, and looking for the distinct combinations of the two (there is usually a 1-1 relationship between the two, but not always)

For smaller recipient types (counties, tribes, and municipalities), about how much have they been awarded?


```{r 2}
library("stringr")
library(scales)
library(gridExtra)

data%>%
  mutate(recipient_amount = paste(Prime.recipient,Award.amount,Sep = " ", callapse = NULL))%>%
  distinct(across(recipient_amount),.keep_all = TRUE)%>%
  summarize(recipients = length(unique(Prime.recipient)))

Distinct_Award_Data <- data%>%
  mutate(Award_number_new = ifelse(Award.number=="",0,Award.number))%>%
  mutate(Award = paste(Prime.recipient, Award_number_new, sep = " - ", collapse = NULL)) %>%
  mutate(Award_amount_detail = ifelse(is.na(Sub.award.amount),Award.amount,Sub.award.amount))%>%
   mutate(Type = if_else(str_detect(Prime.recipient,"STATE|COMMONWEALTH|GOVERNOR"),"State", if_else(str_detect(Prime.recipient,"COUNTY"),"County",if_else(str_detect(Prime.recipient,"TRIBE|NATION|TRIBAL|NATIVE|INDIAN|COUNCIL|COMMUNITY|RANCHERIA"), "Tribe",if_else(str_detect(Prime.recipient,"CITY|MUNICIPALITY|VILLAGE|TOWN|PUBELO"), "City, Town, or Village","Other")))))%>%
   mutate(recipient_amount = paste(Prime.recipient,Award.amount,Sep = " ", callapse = NULL))%>%
  distinct(across(recipient_amount),.keep_all = TRUE)%>%
  select(Prime.recipient, recipient_amount, Award.amount, Type)%>%
    distinct(across(recipient_amount),.keep_all = TRUE) 



PlotA <- Distinct_Award_Data%>%
  filter(Type != "State", Type !="Other")%>%
  ggplot()+
  geom_boxplot(mapping = aes(x = Type, y = Award.amount, na.rm = FALSE))+
  scale_y_continuous(labels = label_number( scale = 1e-9))+
  labs(y = "Award Amount In Billions", x="Award Type")+
  labs(title = "Obvious Outliers in Pandemic Relief Award Recipients", subtitle = "Distribution of pandemic spending amongst localities", caption = "Source: Pandemicoversight.gov")+
  annotate(geom = "text", x = 1.1, y = 1490000000, 
    label = "New York City", hjust = 0, vjust = 1, size = 4)+
    annotate(geom = "text", x = 2.1, y = 1090000000, 
    label = "LA County", hjust = 0, vjust = 1, size = 4)+
  annotate(geom = "text", x = 3.1, y = 750000000, 
    label = "Navajo Nation", hjust = 0, vjust = 1, size = 4)

PlotA

PlotB <- Distinct_Award_Data%>%
  filter(Type == "State")%>%
  ggplot()+
  geom_boxplot(mapping = aes(x = Type, y = Award.amount, na.rm = FALSE))+
  scale_y_continuous(labels = label_number( scale = 1e-9))+
  labs(y="Total Award Amount (Billions)",x=NULL)+
  annotate(geom = "text", x = 1.05, y = 9525564744, 
    label = "California", hjust = 0, vjust = 1, size = 4)+
   annotate(geom = "text", x = 1.05, y = 8000000000, 
    label = "Texas", hjust = 0, vjust = 1, size = 4)

PlotC <- Distinct_Award_Data%>%
  filter(Type == "Other")%>%
  ggplot()+
  geom_boxplot(mapping = aes(x = Type, y = Award.amount, na.rm = FALSE))+
  scale_y_continuous(labels = label_number( scale = 1e-9))+
  labs(y=NULL,x=NULL)+
   annotate(geom = "text", x = 1.05, y = 6000000000, 
    label = "Florida Dept of Financial Services", hjust = 0, vjust = 1, size = 4)

grid.arrange(PlotB, PlotC, ncol=2, top="State and Other Grantees", bottom = "Source: Pandemicoversight.gov")
```



##Visualization Three

Now, let's try to get a better sense at where that money is going, and how much has been spent. I want to narrow in on the sub-recipients, but just for the State, County, City, and Tribal Governments. Let's look at the amount of funding they have actually spent, based on what each government is reporting. Are they doing a good job of reporting their expenditures, and what is it going to? 

```{r 3}

data %>%
  mutate(Type = if_else(str_detect(Prime.recipient,"STATE|COMMONWEALTH|GOVERNOR"),"State", if_else(str_detect(Prime.recipient,"COUNTY"),"County",if_else(str_detect(Prime.recipient,"TRIBE|NATION|TRIBAL|NATIVE|INDIAN|COUNCIL"), "Tribe",if_else(str_detect(Prime.recipient,"CITY|MUNICIPALITY|VILLAGE|TOWN|PUBELO"), "City, Town, or Village","Other")))))%>%
  select(Money.spent.to.date, Spending.category, Type)%>%
  filter(Type !="Other")%>%
  group_by(Spending.category)%>%
    summarize(total = sum(Money.spent.to.date))%>%
  mutate(Category = fct_reorder(Spending.category, total))%>%
  ggplot()+
  geom_col(mapping = aes(x = Category, y = total, fill = total))+
    labs(y = "Total Amount Spent (Billions)")+
  labs(title = "Over $8 billion of unspecified spending from government relief funds",subtitle = "Pandemic Relief Reported Spending",caption = "Source: Pandemicoversight.gov")+
   scale_y_continuous(labels = label_number( scale = 1e-9))+
  theme_minimal()+
  scale_x_discrete(label=function(x) stringr::str_trunc(x, 25))+
  theme(axis.text.x = element_text(angle = 45, hjust=1))

```


##Visualization Four

Finally, let's narrow in on our "Other" recipient type. Can we learn a little bit about where that spending went, if it wasn't directly to states, counties, municipalities, or traibal governments? Most of this funding was labeled as grant funding, so we will focus on just the grant funding that wasn't allocated directly to local governments.

```{r}
library(gridExtra)


Other <- data %>%
   mutate(Type = if_else(str_detect(Prime.recipient,"STATE|COMMONWEALTH|GOVERNOR"),"State", if_else(str_detect(Prime.recipient,"COUNTY"),"County",if_else(str_detect(Prime.recipient,"TRIBE|NATION|TRIBAL|NATIVE|INDIAN|COUNCIL|COMMUNITY|RANCHERIA"), "Tribe",if_else(str_detect(Prime.recipient,"CITY|MUNICIPALITY|VILLAGE|TOWN|PUBELO"), "City, Town, or Village","Other")))))%>%
  filter(Type=="Other")%>%
  filter(Award.type=="GRANT")%>%
  mutate(SUB_Distinct = paste(Prime.recipient,Sub.recipient,Sep = "", callapse = NULL))%>%
  distinct(across(SUB_Distinct),.keep_all = TRUE)%>%
  mutate(Sub_type = if_else(str_detect(SUB_Distinct,"SCHOOL|EDUCATION|ELEMENTARY|ACADEM|CHILD"),"Educational",if_else(str_detect(SUB_Distinct,"HOSPITAL"),"Health Care",if_else(str_detect(SUB_Distinct,"BANK"),"Banks",if_else(str_detect(SUB_Distinct,"MEDIC|HEALTH"),"Health Care",if_else(str_detect(SUB_Distinct,"CHURCH|MINISTRIES|MINISTRY"),"Nonprofit",if_else(str_detect(Award.description,"produce|meat|Meat|Farm|Fish|fish|Produce|Dairy|dairy|Agricultur|agricul|FOOD|food|Food|Nurtition|nutrition|diet"),"Food",if_else(str_detect(Award.description,"School|school|CDL|education|instruction|Education|CHILD|Teacher|learning|Learning|teacher|academic|Academic|child|Child|COLLEGE|College|college|University|UNIVERSITY|university|EDUCATION|LEA"),"Educational",if_else(str_detect(Award.description,"Housing|housing|house|home|HOME|rent|Rent|mortgage|Mortgage|utilit|Utilit"),"Housing",if_else(str_detect(Award.description,"Health Care|Behavioral Health|behavioral health|Mental Health|counseling|MENTAL HEALTH|Local Public Health|mental health|HEALTHCARE|Vaccin|vaccin|VACCIN|testing|Testing|Health Care|HEALTH|Health care|medical|Medical|Hospital|HOSPITAL|hospital|Contact tracing|contact tracing|test|personal protective|PPE|ppe"),"Health Care",if_else(str_detect(Award.description,"Business|business|Tourism|tourism|BUSINESS|TOURISM|Hotel|hotel|To provide short"),"Business",if_else(str_detect(Award.description,"Wireless|wireless|phone|Wifi|wifi|WIFI|WiFi|Technical assistance|technical assistance|Technical Assistance|broadband|Broadband|hotspot|Hotspot|digital|Digital|remote|Technology|technology|Fiber|fiber"),"Technology",if_else(str_detect(Award.description,"fuel|Fuel|FUEL|Ethanol|ethanol"),"Fuel",if_else(str_detect(Award.description,"CHURCH|MINISTRIES|MINISTRY|Religio|religio"),"Nonprofit",if_else(str_detect(Award.description,"Nonprofit|nonprofit|Non-profit|non-profit|Community based|(COA)|Community investments|Make investments in communities to improve|Community Assistance|Community assistance|community assistance|victim|Victim|REFUGEE"),"Nonprofit",if_else(str_detect(Award.description,"music|art|culture|entertainment|Music|Art|Culture|Entertainment"),"Arts",if_else(str_detect(Award.description,"reimbursement|Reimbursement|Municipalities|municipalities|local economic support|Distribution to local|MITIGATION|MUNICIPALITIES|COUNTIES|Counties|counties|CRF"),"Local governments",if_else(str_detect(Award.description,"employment|Employment"),"Employment",if_else(str_detect(Award.description,"Payroll|payroll|Hazard Pay|Hazard pay|hazard pay"),"Payroll","Other")))))))))))))))))))%>%
  select(SUB_Distinct, Sub.award.amount, Sub_type, State, Type, Award.type, Award.description)

OtherOther<- Other%>%
  filter(Sub_type=="Other")

Plot1<-Other %>%
  group_by(Sub_type)%>%
  summarize(total = sum(Sub.award.amount))%>%
  ggplot()+
   geom_col(mapping= aes(x = Sub_type, y=total))+
   scale_y_continuous(labels = label_number( scale = 1e-9))+
   theme(axis.text.x = element_text(angle = 45, hjust=1))+
  labs(y = "Total Awarded (Billions)")+
  labs(x = "Award Category")



Plot2 <- Other%>%
  group_by(Sub_type)%>%
  summarize(count = length(SUB_Distinct))%>%
  ggplot()+
  geom_point(mapping = aes(x=Sub_type, y = count))+
  theme(axis.text.x = element_text(angle = 45, hjust=1))+
   labs(y = "Total Recipients")+
  labs(x = "Award Category")


grid.arrange(Plot1, Plot2, ncol=2, top="Breakdown of Grants Awarded for Covid Relief Funding", bottom = "Source: Pandemicoversight.gov")



```
##Conclusion

This visualization exercise taught me more about the shortcomings of the dataset than about government expenditures for covid relief. Although this data has been made public, there is little consistency in the grantees expectation to report expenditure, nor are the awards labeled with clear descriptors in order to ensure effective oversight. Some helpful categories and labels on the data would have been dates of funding, dates of expenditure, Clear Grant categories, and categories for sub-recipients, such as nonprofits, business, healthcare institutions, etc. 
